---
title: "Human Resource Analytics"
output:
  html_document: default
  pdf_document: default
---


---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Loading The required packages
```{r  , message=FALSE,warning=FALSE}
library(ggplot2)
library(dplyr)

```


```{r,message=FALSE,warning=FALSE}
#loading the Dataset

hrm<-read.csv('HR_comma_sep.csv')

#Structure of the Dataset
str(hrm)
attach(hrm)

#converting left variable to factor variable 
hrm$left<-ifelse(left==1,'True','False')

hrm$left<-factor(hrm$left,levels=c("True","False"))
table(hrm$left)

#Summary Statistics of the dataset
summary(hrm)




```


---

<h3>Satisfaction level statistics splitted by salary ranges</h3>
```{r}

by(hrm$satisfaction_level,hrm$salary,summary)

#Histogram

p1<-ggplot(aes(x=satisfaction_level),data=hrm) + 
  geom_histogram(color="black",fill="red",bins = 30) +
  labs(title="Satisfaction level Histogram",x='Satisfaction Level of Employees', y="Frequency")

p1



    
#Satisfaction level histogram facetted by sallary classes

p2 =  p1  + facet_wrap(~salary)
p2

```

---

<ul style="font-size:18px">
<li> The disrtibution of satisfaction level for each class of Salary Ranges is almost same.</li>
<li> The High salary employees have a little bit higher mean satisfaction level.</li>
<li>
Employees with High salary are less in number.
</li>
</ul>

```{r}
by(satisfaction_level,left,summary)
#As peedicted the satifaction level of employees who left was lower

#Sstisfaction level vs left
ggplot(aes(x = satisfaction_level),data=hrm) + 
  geom_histogram(color='black',fill='green',bins=35) +
  xlab('Satisfaction Level') + 
  ylab("Frequency")  + 
  facet_wrap(~left)
  

#Boxplot for Satisfaction level vs left
ggplot(aes(x = left,y=satisfaction_level),data= hrm) + 
  geom_boxplot() + 
  ylab('Satisfaction Level') + 
  xlab("Employee left") + 
  labs(fill="Salary Classes")

#Boxplot for Satisfaction level vs left facetted by Salary Ranges
ggplot(aes(x = left,y=satisfaction_level),data= hrm) + 
  geom_boxplot() + 
  ylab('Satisfaction Level') + 
  xlab("Employee left") + 
  facet_wrap(~salary)

table(hrm$left , salary)

#Testing for the dependence between left and salary Ranges
#Both are categorial variables so we use Chisq Test statistic
chisq.test(left,salary)




```
### X-squared value is high and p-value is less i.e results are significant.Both variables are related

---

## Analysis on number of Projects



```{r}
hrm$number_project<-factor(hrm$number_project)

ggplot(aes(x=number_project),data = hrm) + 
  geom_bar(color='black',fill='#234338') +
  xlab("Number of Projects") + 
  ylab("Frequency") + 
  labs(title="Barplot of Number of projects")





#boxplot of number of projects vs  Average monthly hours at workplace of employees
p3=ggplot(aes(x=number_project, y = average_montly_hours),data=hrm)+
  geom_boxplot()

p4=p3+facet_wrap(~salary)
p4

p5=p3+facet_wrap(~left) + labs(title="Number projects Vs Avg monthly hours worked faceted by Left")
p5


#facetted by salary
ggplot(aes(x=number_project),data = hrm) + 
  geom_bar(color='black',fill='#834338') +
  xlab("Number of Projects") + 
  ylab("Frequency") + 
  labs(title="Barplot of Number of projects faceted by Salary") +
  facet_wrap(~salary)


#faceted by If a employee left or not
ggplot(aes(x=number_project),data = hrm) + 
  geom_bar(color='black',fill='#547398') +
  xlab("Number of Projects") + 
  ylab("Frequency") + 
  labs(title="Barplot of Number of projects faceted by Left")+  
  facet_wrap(~left)
  

```

---

## Analysis on Average Number of Hours a Employee works

```{r}
#Analysis of average monthly hours
summary(average_montly_hours)
#Somewhat Normally distributed

ggplot(aes(x= average_montly_hours),data = hrm)+
  geom_histogram(color='black',fill="yellow",bins = 30)


cor.test(satisfaction_level,average_montly_hours)
#No relation between both the variables -  as r is eqv to 0

```



```{r}

ggplot(aes(x = average_montly_hours),data =hrm ) + 
  geom_histogram(color='black',fill='#443332',bins = 30) + 
  facet_wrap(~left)

by(average_montly_hours , hrm$left ,summary)

ggplot(aes(y = average_montly_hours, x = hrm$left),data=hrm)+
  geom_boxplot() + 
  xlab("Employee left or not") + 
  ylab("Average Montly hours worked")



```

### A thing to notice is that employee who left the company worked more hours than those who did not leave, hence it might be possible that they left bacause they were over pressurized by their peers or bosses or over worked or stressed with lots of work




## Anslysis for variable Time spend at company
```{r}
table(hrm$time_spend_company)

ggplot(aes(x = factor(time_spend_company)),data = hrm) + 
  geom_bar(fill = 'purple',color='black') + 
  xlab("Time spend at compnay in years") + 
  ylab("Frequency")+
  labs(title = "Barplot of Time spend at Company")


#Time spend at company vs Left or not

ggplot(aes(x = factor(time_spend_company)),data = hrm) + 
  geom_bar(fill = 'grey',color='black') + 
  xlab("Time spend at compnay in years") + 
  ylab("Frequency")+
  labs(title = "Barplot of Time spend at Company faceted by Left")  +
  facet_wrap(~left)

by(time_spend_company , left , summary)

ggplot(aes(x = left , y = time_spend_company),data = hrm)+
  geom_boxplot()



```

### Time Spend at company vs Satisfaaction level
```{r}

by(satisfaction_level,factor(time_spend_company),summary)

cor.test(satisfaction_level,time_spend_company)
#both have a negetive correlation

#plots vs Time spend and Satisfaction level
ggplot(aes(x=factor(time_spend_company),y=satisfaction_level),data=hrm)+
  geom_boxplot() + 
  xlab("Time spend at company in years")+ 
  ylab("Satisfaction level")





#Time spend at compnay vs Promotion in last 5 years

table(Promotion=promotion_last_5years,Time_Spend=factor(time_spend_company))
#Employees who have had promotion are very less


ggplot(aes(x = factor(time_spend_company)),data = hrm)+
  geom_bar()+
  facet_wrap(~promotion_last_5years) + 
  scale_y_continuous(limits=c(0,4000),breaks=seq(0,4000,500))





#Time spend vs Department of Work

by(time_spend_company,sales,summary)


ggplot(aes(x =sales),data = hrm ) +
  geom_bar()  +
  xlab('Department') + 
  ylab('Counts') +
  coord_flip() 
#highest count is for Sales department then Technical  and least for 
#Management

ggplot(aes(x = sales,y = time_spend_company),data = hrm) + 
  geom_boxplot() + 
  coord_flip()









```


---



## Analysis of Department of Work 

```{r}

ggplot(aes(x =sales),data = hrm ) +
  geom_bar()  +
  xlab('Department') + 
  ylab('Counts') +
  coord_flip() 
#highest count is for Sales department then Technical  and least for 
#Management


#Department vs sallary

table(Dept = sales , Salary  = salary)


ggplot(aes(x =sales),data = hrm ) +
  geom_bar(aes(fill=salary))  +
  xlab('Department') + 
  ylab('Counts') +
  coord_flip()


ggplot(aes(x =sales),data = hrm ) +
  geom_bar()  +
  xlab('Department') + 
  ylab('Counts') +
  labs(title = "Department and their count facetted by Salary ranges")+
  facet_wrap(~salary) + 
  coord_flip()

chisq.test(sales,salary)
#Department and Salary is dependent on each other . 




#finding proportions
prop.table(table(Dept = sales , left = left))*100

as.data.frame(table(sales , left))->deptdf
deptdf


deptdf<-hrm %>% group_by(sales,left) %>% 
      summarise(count=n())

require(tidyr)
#making a dat frame of Departments and the count of workers who left or not
deptdf<-spread(deptdf,left,count)

deptdf<-transform(deptdf,Perleft=(True/(True+False))*100 , PerWork=(False/(True+False))*100)
deptdf



#Plot of Department vs Percentage of Employees who left
ggplot(aes(x=sales, y = Perleft),data = deptdf) + 
  geom_col(fill='#53ab85',color='#2f3f52') + 
  coord_flip()+
  xlab("Department") + 
  ylab("Percentage of Employees who left") + 
  labs(title="Plot of Department vs Percentage of Employee left")
#highest percentage of employees belonged to HR dept then accounting
# least for management dept who left





#Plot of Department vs Percentage of People Working
ggplot(aes(x=sales, y = PerWork),data = deptdf) + 
  geom_col(fill='#b6a2bf',color='#2f3f52') + 
  coord_flip()+
  xlab("Department") + 
  ylab("Percentage of Employees who Still Work") + 
  labs(title="Plot of Department vs Percentage of Employees Working")





#Department vs Satisfaction level

by(satisfaction_level,sales,summary)
#highest mean satisfaction for R&D and Management Dept

ggplot(aes(x = sales, y = satisfaction_level),data = hrm)+
  geom_boxplot() + 
  scale_y_sqrt()+
  xlab('Department') + 
  ylab('Satisfaction Level"') + 
  coord_flip()
#Highest Median Satisfaction for IT dept, R&D and , Management
#Least Median Satifaction level for HR and Accounting







#Analysis of Department vs Time spend at company

by(time_spend_company,sales,summary)

#Maximum  Mean Time spent by Managaement Employees

ggplot(aes(x = sales,y = time_spend_company),data = hrm) + 
  geom_boxplot() +
  xlab('Department') + 
  ylab("Time Spend at Company") + 
  coord_flip()


ggplot(aes(x = time_spend_company),data = hrm) + 
  geom_bar() +
   xlab("Time Spend at Company splitted by Department") + 
facet_wrap(~sales)
#In every department there is very less count of Employees
# working for over 5 years






#Department vs Time average monthly hours

by(average_montly_hours,sales , summary)


#Highest average working time for IT and Technical departments

ggplot(aes(x = sales , y = average_montly_hours),data =hrm) +
  geom_boxplot() + 
  xlab('Department of Work') + 
  ylab('Average Monthly Hourse of Work') + 
  coord_flip()

#Highest Median working time of Management department





#Department vs Work Accident

table(Work_accident)

table(sales,Work_accident)

ggplot(aes(x = sales),data = hrm) +
  geom_bar(aes(fill=factor(Work_accident))) + 
  coord_flip() + 
  labs(x = "Department",y ="Frequency", fill="Work Accidents" )


hrm$Work_accident<-factor(Work_accident,labels = c('False','True'))

accidentdf<-hrm %>% group_by(sales,Work_accident) %>%
  summarise(Count= n())

accidentdf<-spread(accidentdf,Work_accident,Count)

accidentdf<-transform(accidentdf,TrueRate=(True/(True+False))*100,FasleRate=(False/(True+False))*100)

#Plot of Departent vs Accidental Rate 
ggplot(aes(x = sales,y = TrueRate),data = accidentdf) + 
  geom_col(color='black',fill="#b266b2") + 
  xlab('Department') + 
  ylab('Accident Percentage') + 
  coord_flip()
#Hishest number of accidents in R and D department



ggplot(aes(x = sales,y = FalseRate),data = accidentdf) + 
  geom_col(color='black',fill="#d8b2d8") + 
  xlab('Department') + 
  ylab('No Accident Percentage') + 
  coord_flip()
#Maximum for HR department





#Department vs number_projects made

by(number_project,sales,summary)

ggplot(aes(x = sales, y =factor(number_project)),data = hrm) +
  geom_count() +
  xlab("Department") + 
  ylab("Number of projects") + 
  labs(title = "Plot of Department vs Number of projects and their count ")





#Department vs Promotion in last 5 years

table(sales , hrm$promotion_last_5years)

#TRansforming Promotion Column to Factor with True and False values
hrm$promotion_last_5years<-factor(promotion_last_5years,labels=c('False',"True"))

#Generating a promotions Data frame
promotiondf<-hrm %>% group_by(sales,promotion_last_5years) %>%
  summarise(Count = n())

#Sprading the data'
promotiondf<-promotiondf %>% spread(promotion_last_5years,Count)

#changing the names
names(promotiondf)<-c("Department","No Promotion","Promotion")

#replacing NA valuw with 0
promotiondf[is.na(promotiondf)]<-0


```

